### 2013-10-23 plan.tex
Started an outline and added a single reference.

\section{Plan}

We propose to create a verified compiler for a language similar to that
discussed in Pottier and Simonet's work\cite{InfoFlowML}.  We will
construct this compiler using the Coq proof assistant\cite{coq?} using
features which allow us to extract a certified compiler.  This will
allow us to take the first step towards a verified compiler for a
language with static information flow types such as Jif\cite{Jif?} or
FlowCaml\cite{FlowCaml?}.

%TODO:
Stuff about why it would be useful to have a verified compiler for
something like Jif or FlowCaml.

%TODO:
Stuff about the target and source languages we will be supporting.

%TODO:
Expected timeline, anticipated problems, and potential additions if we
have time?