### 2013-10-24 plan.tex
Tweaks.

\section{Plan}

We propose to create a verified compiler for a language similar to that
discussed in Pottier and Simonet's work\cite{InfoFlowML} compiling to a
simple register transfer level (RTL) language.  We will construct this
compiler using the Coq proof assistant\cite{coq?} using features which
allow us to extract a certified compiler.  This will allow us to take
the first step towards a verified compiler for a language with static
information flow types such as Jif\cite{myers1999jflow} or
FlowCaml\cite{simonet2003flow,simonet2003flowmanual}.

%TODO:
Stuff about why it would be useful to have a verified compiler for
something like Jif or FlowCaml.

%TODO:
Expected timeline, anticipated problems, and potential additions if we
have time?