### 2013-10-24 plan.tex
Draft of plan.  Added some references.

\section{Plan}

We propose to create a verified compiler for a language similar to that
discussed in Pottier and Simonet's work\cite{InfoFlowML} compiling to a
simple register transfer level (RTL) language.  We will construct this
compiler using the Coq proof assistant\cite{coqart} using features which
allow us to extract a certified compiler.

This will allow us to take the first step towards a verified compiler
for a language with static information flow types such as
Jif\cite{myers1999jflow} or
FlowCaml\cite{simonet2003flow,simonet2003flowmanual}.  Developing a
verified compiler for such languages removes the need for programmers to
trust the compiler, they can instead verify the proofs used to generate
the compiler to convince themselves that semantics of the original
source program are preserved and properties such as non-interference are
enforced by the compiler.  This rules out the possibility that a
compiler does not introduce bugs not present in the source program, such
as additional undesirable information flows.  As a result, this compiler
reduces the trusted computing base from a large complex compiler to a
much simpler verifier to mechanically check the proofs the compiler is
extracted from.

To develop this verified compiler using Coq, we plan to write formal
proofs to show that the type system correctly enforces non-interference
and that a CPS conversion\cite{appel2007compiling}, which would be our
approach to compiling the language to an RTL language, correctly
preserves the semantics of the source program.  We expect that the proof
that the type system enforces non-interference will be very similar to
the proof presented in Pottier and Simonet's paper\cite{InfoFlowML}.
Development will proceed by first writing the proof for CPS conversion,
developing a compiler for the language without enforcing the type
system, and then going back and adding the non-interference proof.